Modeling the near-wake of a vertical-axis cross-flow turbine with 2-D and 3-D RANS

### 2015-09-14 paper.tex
A word

In the pursuit of a fully sustainable energy profile, cross-flow (often
vertical-axis) turbines (CFTs) still play a role. Despite extensive research and
development in the 1970s through the early 1990s by groups like Sandia National
Labs in the US \cite{Sutherland2012} and the National Research Council in Canada
\cite{Para2002}, the vertical-axis cross-flow Darrieus turbines were all but
abandoned for large scale commercial terrestrial wind power. Today, however,
with the development of marine hydrokinetic (MHK) energy devices, CFTs are at
the forefront, with the first grid-connected commercial device being a
horizontal-axis CFT manufactured and installed by ORPC in Cobscook Bay, Maine
\cite{ORPC2012}. CFTs are also being examined for small-scale onshore wind in
urban areas \cite{Lott2015}, offshore floating wind farms \cite{Paulsen2011,
    Sandia2012}, and onshore wind farms where power output per unit land area is of
interest \cite{Dabiri2011}.

>>>>

In the pursuit of a fully sustainable energy profile, cross-flow (often
vertical-axis) turbines (CFTs) can still play a role. Despite extensive research
and development in the 1970s through the early 1990s by groups like Sandia
National Labs in the US \cite{Sutherland2012} and the National Research Council
in Canada \cite{Para2002}, the vertical-axis cross-flow Darrieus turbines were
all but abandoned for large scale commercial terrestrial wind power. Today,
however, with the development of marine hydrokinetic (MHK) energy devices, CFTs
are at the forefront, with the first grid-connected commercial device being a
horizontal-axis CFT manufactured and installed by ORPC in Cobscook Bay, Maine
\cite{ORPC2012}. CFTs are also being examined for small-scale onshore wind in
urban areas \cite{Lott2015}, offshore floating wind farms \cite{Paulsen2011,
    Sandia2012}, and onshore wind farms where power output per unit land area is of
interest \cite{Dabiri2011}.
